Supplementary Materials

### 2023-06-09 manuscript/manuscript.tex
MS updates, figs

@@

\subsection*{DFE Estimates}
% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA

@@

\section*{Methods}

>>>>

\subsection*{Annotation Model Comparison}

Our method takes tracks of annotated features (an ``annotation model") that are
\emph{a priori} expected to have similar fitness effects, and estimates the
distribution of fitness effects for each feature type. We consider two classes
of annotation models: (1) CADD score models, which assign a rank-order
pathogenicity score to each basepair, and (2) and more interpretable gene
feature-based models that includes protein coding regions, intron and UTRs, and
PhastCons regions. We include PhastCons regions because there is abundant
evidence of highly-conserved non-coding regions in humans that we would expect
to be under negative selection and impact patterns of linked selection (XXX).
Since our method fits a DFE for each feature type, PhastCons regions that
overlap other gene annotations (e.g. coding sequences) must be assigned to
either of the two feature types. Since this can impact estimates, we fit both
alternatives: a PhastCons Priority model, where features that overlap PhastCons
regions are classified as PhastCons, and a Feature Priority model, where gene
structure features take priority and the PhastCons type catches all
highly-conserved non-genic regions. In total, we fit four annotation models
(CADD 6\%, CADD 8\%, PhastCons Priority, and Feature Priority) to high-coverage
1000 Genome data for three populations: Yoruba (YRI), Han Chinese (CHB), and
European (CEU) ancestries.

Overall, we find the PhastCons Priority and CADD 6\% models fit about equally
as well (Figure \ref{fig:figure-2}A), explaining $R^2=68.45$\% and
$R^2=68.04$\% of the out-sample variance in Yoruba pairwise diversity at the
megabase-scale, respectively. In populations impacted by the out-of-Africa
bottleneck, the goodness-of-fit was lower (62.98\% and 59.98\% for CEU and CHB,
respectively). The CADD 6\% model was previously found by
\textcite{Murphy2022-sj} to be the best-fitting model using classic BGS. Our
PhastCons Priority model fits marginally better than the CADD 6\% model
consistently across all three populations (Supplementary Table XXX). This is
likely because the PhastCons Priority model has more DFE parameters, and thus
can fit finer-scale differences in the DFE across its four feature types. By
contrast, the CADD 6\% track must fit patterns of diversity to conserved sites
under a single DFE. The CADD 8\% model fit slightly worse than the CADD 6\%
across all populations, and the Feature Priority model had the poorest-fit.

Since our method is built upon theory that fixes the weak selection problem of
classic BGS theory, it should in principle fit equally well when the annotation
model covers the full genome, the bulk of which is expected to be neutrally
evolution. To test this, for each annotation model (which is ``sparse") we fit
a ``full" model that contains the remaining complement of the genome labeled as
``other". Ideally, the goodness-of-fit of fit using our B' method should be
invariant to whether an annotation model is sparse or full. Indeed, we find
that

NOTE: there looks like a bug in the table... repeated values.

Each model was fit to the high-coverage 1000 human genome data for
three populations, YRI, CEU, and CHB. Additionally, we also fit each of these
models using our method, but using classic BGS theory rather than our B'
theory. In total, we fit and compared 48 different models.

Overall, we find that the goodness-of-fit, measured by both in- and out-sample
$R^2$ are nearly identical across sparse and full fits using our B' method,
regardless of the different annotation models (Figure \ref{fig:figure-2},
Supplementary Materials Table XXX). By contrast, we find that when set our
method to use classic BGS theory, the model fits are worse in full-coverage
annotation models (Figure \ref{fig:figure-2}B). Additionally, when sparse
annotation models contain genomic features that are likely under weak
constraint (such as introns and UTRs), models fit worse under classic BGS
theory than our B' method (Figure \ref{fig:figure-2}A). However, among the CADD
annotation models, the goodness-of-fit is nearly identical between B' and
classic BGS methods. This behavior is what we would expect given that the CADD
annotation models only contain the most pathogenic sites, which are \emph{a
prior} very likely under the strong selection domain under which B' and classic
BGS theory agree.

The out-sample $R^2$

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all 
    populations (colors) fit at the megabase-scale. Round points are our B' 
    method; diamonds are the classic B. For our method, we also calculate 
    leave-one-out out-sample $R^2$ (lighter filled points).}
  \label{fig:figure-2}
\end{figure}

@@

\subsection*{DFE Estimates}

Our best-fitting models have out-sample $R^2$ of around 68\% for Yoruba, 62\%
European, and 60\% for Han Chinese ancestry individuals. However, $R^2$ has a
theoretic upper bound due to coalescence noise around the predicted mean level
$\widehat{\pi}(x)$.

% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA

@@

\section*{Methods}
% (the method would support a new
% ``interaction" type, but this greatly increases the number of parameters).

### 2023-06-09 manuscript/supplementary.tex
MS updates, figs

@@

where $w$ is the width of the window such that the per-basepair $\pi$ is
appropriately scaled. Then, we note that the variance in pairwise diversity is
equivalent to the variance in the number of segregating sites for a sample of 
two, $\var(\pi) = \var(S_2)$. From \textcite[eq. 7.20]{Wakeley2009-ua}

@@

Since this is a rough approximation, we set $\theta = w \pi_0 B_i$ and fix
$\rho = w \gamma \pi_0 B_i$ where $\gamma$ is the ratio of recombination to
mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but we
wanted to check the sensitivity of this calculation to this.

This back-of-the-envelope calculation finds that the upper bound of $R^2$ would
be around 74\% for Yoruba, and 80-81\% for European and Han Chinese models
(Figure \ref{suppfig:r2_upper} We note, however, that this rough approximation
assumes constant demography, and that the out-of-Africa bottleneck causes a
reduction in effective population size that would shrink the variance of
exponential coalescence times, but a bottleneck could in fact inflate these
variances. A more accurate approximate of the upper $R^2$ bound could rely on
simulations with more realistic demography.

>>>>

\section*{Table of $R^2$ Values}

\begin{table}
\centering
\caption{$R^2$ and mutation rate estimates for all models}
\begin{tabular}{lll|crr|cc}
    \textbf{Model} & \textbf{Track type} & \textbf{Pop.} & \textbf{B' $R_\text{LOO}^2$} & \textbf{B' $R_\text{IS}^2$} & \textbf{B $R_\text{IS}^2$} & \textbf{B' $\hat{\mu} \times 10^{-8}$} & \textbf{B $\hat{\mu} \times 10^{-8}$} \\[0.5ex] 
\hline
\hline
    \textbf{phastcons$>$CDS$>$genes} &            sparse &          YRI &                        \textbf{68.45} &             68.16 &            65.23 &                                 1.8636 &                                0.2975 \\
phastcons$>$CDS$>$genes &              full &          YRI &                        68.19 &             68.12 &            63.67 &                                 1.7082 &                                0.1666 \\
               CADD 6\% &              full &          YRI &                        68.08 &             68.08 &            62.96 &                                 1.4612 &                                0.1721 \\
               CADD 6\% &            sparse &          YRI &                        68.04 &             68.03 &            68.01 &                                 2.0912 &                                2.0716 \\
               CADD 8\% &              full &          YRI &                        66.98 &             67.11 &            63.61 &                                 1.0126 &                                0.1713 \\
               CADD 8\% &            sparse &          YRI &                        66.67 &             67.01 &            66.98 &                                 1.5688 &                                1.5473 \\
CDS$>$genes$>$phastcons &              full &          YRI &                        64.90 &             65.51 &            61.23 &                                 4.1951 &                                0.1620 \\
CDS$>$genes$>$phastcons &            sparse &          YRI &                        64.90 &             65.51 &            63.16 &                                 4.1795 &                                0.3070 \\
\textbf{phastcons$>$CDS$>$genes} &            sparse &          CEU &                        \textbf{61.98} &             63.02 &            60.02 &                                 2.0194 &                                0.3036 \\
phastcons$>$CDS$>$genes &              full &          CEU &                        61.78 &             63.01 &            58.57 &                                 2.3499 &                                0.1679 \\
               CADD 6\% &              full &          CEU &                        61.74 &             62.94 &            57.42 &                                 1.5365 &                                0.1739 \\
               CADD 6\% &            sparse &          CEU &                        61.66 &             62.86 &            62.84 &                                 2.1215 &                                2.1084 \\
               CADD 8\% &              full &          CEU &                        60.80 &             62.12 &            57.95 &                                 1.1572 &                                0.1742 \\
               CADD 8\% &            sparse &          CEU &                        60.59 &             62.06 &            62.04 &                                 1.6030 &                                1.5876 \\
               CADD 6\% &              full &          CHB &                        59.62 &             61.31 &            55.31 &                                 1.5083 &                                0.1727 \\
               CADD 6\% &            sparse &          CHB &                        59.54 &             61.21 &            61.20 &                                 2.1242 &                                2.1170 \\
               \textbf{phastcons$>$CDS$>$genes} &            sparse &          CHB &                        \textbf{59.44} &             61.08 &            57.88 &                                 2.1982 &                                0.3060 \\
phastcons$>$CDS$>$genes &              full &          CHB &                        59.29 &             61.08 &            56.71 &                                 2.5573 &                                0.1694 \\
               CADD 8\% &              full &          CHB &                        58.88 &             60.63 &            55.86 &                                 1.1128 &                                0.1733 \\
               CADD 8\% &            sparse &          CHB &                        58.65 &             60.57 &            60.55 &                                 1.6098 &                                1.6001 \\
CDS$>$genes$>$phastcons &            sparse &          CEU &                        58.58 &             60.30 &            58.41 &                                 4.1041 &                                0.3110 \\
CDS$>$genes$>$phastcons &              full &          CEU &                        58.58 &             60.30 &            56.75 &                                 4.1001 &                                0.1648 \\
CDS$>$genes$>$phastcons &            sparse &          CHB &                        56.48 &             58.73 &            56.60 &                                 3.7195 &                                0.3136 \\
CDS$>$genes$>$phastcons &              full &          CHB &                        56.48 &             58.73 &            54.95 &                                 3.7552 &                                0.1654 \\
\hline
\end{tabular}
\end{table}

@@

where $w$ is the width of the window such that the per-basepair $\pi$ is
appropriately scaled. We use the predicted $\hat{B}_i$ as the plugin estimator
for $B_i$. Then, we note that the variance in pairwise diversity is equivalent
to the variance in the number of segregating sites for a sample of two,
$\var(\pi) = \var(S_2)$. From \textcite[eq. 7.20]{Wakeley2009-ua}

@@

Since this is a rough approximation, we set $\theta = w \pi_0 B_i$ and fix
$\rho = w \gamma \pi_0 B_i$ where $\gamma$ is the ratio of recombination to
mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but explore
different ratios to assess sensitivity of this calculation (Supplementary
Figure \ref{suppfig:r2_upper}).

This back-of-the-envelope calculation finds that the upper bound of $R^2$ would
be around 74\% for Yoruba, and 80-81\% for European and Han Chinese models
(Figure \ref{suppfig:r2_upper} This rough approximation assumes constant
demography, which for bottlenecked out-of-Africa populations assumes that the
variance in coalescence times is determined entirely by reducing the effective
population size. However, the variance in coalescence times is \emph{higher} in
bottlenecked populations (CEU and CHB) compared to Yoruba, which we confirm
with simple simulations using the \texttt{OutOfAfrica\_3G09} model from
\texttt{stdpopsim} \parencite{Gutenkunst2009-pg,Adrion2020-cf}. This higher
than expected variance would inflate $SS_\text{res}$, which would decrease

A more accurate approximate of the upper $R^2$ bound could rely on simulations
with more realistic demography.
